Counting infinitely by Oritatami co-transcriptional folding

### 2019-08-23 paper.tex
writting zig...

%\titlerunning{}
\author{
Kohei Maruyama\inst{1}\thanks{Primary corresponding author (\email{k.maruyama@uec.ac.jp})} \and
Shinnosuke Seki\inst{1,2}\thanks{Secondary corresponding author (\email{s.seki@uec.ac.jp})}
}
\institute{
The University of Electro-Communications, 
%Graduate School of Informatics and Engineering, 
1-5-1 Chofugaoka, Chofu, Tokyo, 1828585, Japan \and
ENS Lyon
}

@@

A fixed bit-width (finite) binary counter has been implemented as a proof-of-concept demonstration of oritatami model of cotranscriptional folding \cite{GeMeScSe2019}. 
As shown in Fig.~\ref{fig:rna_origami}, an RNA transcript folds upon itself while being transcribed (synthesized) from its corresponding DNA template strand. 
Geary et al. programmed a specific RNA rectangular tile structure into a DNA template in such a way that the corresponding RNA transcript \textit{folds cotranscriptionally} into the programmed tile structure highly likely \textit{in vitro} at room temperatures (\textit{RNA origami}) \cite{GearyRothemundAndersen2014}. 
An oritatami system folds a transcript of abstract molecules called \textit{beads} of finitely many types cotranscriptionally over the 2-dimensional triangular lattice cotranscriptionally according to a rule set that specifies which types of molecules are allowed to bind. 
The transcript of the binary counter in \cite{GeMeScSe2019} is of period 60 as $\textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1}{-}\textcircled{\scriptsize 2}{-} \cdots {-}\textcircled{\scriptsize 58}{-}\textcircled{\scriptsize 59}{-}\textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1} \cdots$ and its period is semantically divided into two half-adder (HA) modules $A = \textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1}{-} \cdots {-}\textcircled{\scriptsize 11}$ and $C = \textcircled{\scriptsize 30}{-}\textcircled{\scriptsize 31}{-} \cdots {-}\textcircled{\scriptsize 41}$ and two structural modules $B$ and $D$, which are sandwiched by half-adder modules.
While being folded cotranscriptionally in zigzags, HA modules increment the current count $i$ by 1, which is initialized on a linear \textit{seed} structure, alike the Evans' counter, whereas structural modules $B$ and $D$ align HA modules properly and also make a turn at an end of the count $i$; $B$ guides the transcript from a zig to a zag ($\hookrightarrow$) while $D$ does from a zag to a zig ($\hookleftarrow$). 
This counter was embedded as a component of an oritatami system to self-assemble an arbitrary finite portion of Heighway dragon fractal \cite{MasudaSekiUbukata2018}. 
Its applications is limited, however, by lack of mechanism to widen bit-width; its behavior is undefined when its count overflows. 
In this paper, we endow this counter, or more precisely, its structural module $B$, with capability to widen the count by 1 bit at every encounter with overflow.

@@

@@

\subsection{General idea}
Between two consecutive overflows, the proposed system behaves in the same way as the finite binary counter of Geary et al. \cite{GeMeScSe2019}.
Its transcript folds (downward in figures throughout this paper) in a zigzag manner macroscopically.
A zig, folding from right to left, increments the current value of the counter by 1.
The succeeding zag, folding from left to right, formats the incremented value for the sake of next zig and copies it downward.
Unlike the counter of Geary et al., when a zig encounters overflow, it does not abort but extends the current value by 1 bit.

@@

\subsubsection{Encoding}
The binary representation of the initial value, $b_{k-1} b_{k-2} \cdots b_1 b_0$ is encoded on the seed in the following format
\begin{equation} \label{eq:zagencoding}
64{-}65{-}66{-}\left( \prod^0_{i = k-1} \bigl(  w_{Hn} w_{Rb} w_{Fb_i} w_{Lbn} \bigr) \right) w_{Hn}
\end{equation}
where $w_{Hn}, w_{Rb}, w_{Fb_i}, w_{Lbn}$ are sequences of bead types exposed downward by modules H, R, F, L when they fold into bricks Hn, Rb, F$b_i$, Lbn, respectively; these bricks can be found in Figs.~\ref{fig:halfadders},\ref{fig:rightturns}, \ref{fig:formatters}, and \ref{fig:leftturns}.
%$w_{Hn} = 67{-}76{-}77{-}78{-}79{-}88{-}89{-}90{-}91{-}96$ \\
%$w_{Rb} = 97{-}102{-}103{-}108{-}109{-}114{-}115{-}120{-}121{-}126{-}127{-}132$\\
%$w_{Lbn} = 31{-}36{-}37{-}42{-}43{-}48{-}49{-}54{-}55{-}64{-}65{-}66$

%64 65 66
%67 76 77 78 79 88 89 90 91 96
%97 102 103 108 109 114 115 120 121 126 127 132
%
%1 10 11 12 13 22 23 24 25 30
%
%31 36 37 42 43 48 49 54 55 64 65 66
%67 76 77 78 79 88 89 90 91 96
%97 108 109 120

The current value of the counter is encoded in standard binary with the most significant bit to the left.
Binary is encoded into a specific folding of the module F and H.
In zig pass, folding of module H corresponds to module F in a zag pass: namely H0 for 0, and H1 for 1.
In addition, start position of modules represents the carry: carry = 0 if module H starts to fold in the top row; carry = 1 if module H start to fold from the bottom row.
On the other hand in zag pass, folding of module F corresponding to module H in a zig pass: namely F0 for 0, and F1 for 1.

\subsubsection{In the zig pass ($\leftarrow$)}
Module H read from the row above the value encoded into folding in the module F during the previous zag-phase, and fold into a shape H00, H10, H01, H11.
H$xc$ is corresponding to the case where $x$ is the bit read in the row above and $c$ is the carry. 
In the zig pass module R, F, and L just propagate the carry value to the next.

@@

\subsubsection{In the zag pass ($\rightarrow$)}
Module F read form the row above H$xc$, and fold into a shape F0, F1.
F$y$ is corresponding to the case where $y = (x + c)$ mod $2$.
There are no carry propagation and the module L, H, R fold into Lbc, Hn, Rb in this pass.

@@

\end{itemize}
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/svg/CounterEx5_1.pdf}
\caption{
The first zig pass.
The seed is encoded with "0" same as bottom of F0(
\textcircled{\scriptsize 1}{-}\textcircled{\scriptsize 10}{-}\textcircled{\scriptsize 11}{-}\textcircled{\scriptsize 12}{-}\textcircled{\scriptsize 13}{-}\textcircled{\scriptsize 22}{-}\textcircled{\scriptsize 23}{-}\textcircled{\scriptsize 24}{-}\textcircled{\scriptsize 25}{-}\textcircled{\scriptsize 30}),
 and the counter increments it by giving a carry.
Module H outputs "1" as a sum and cancels the carry.
}
\label{fig:counter1stzig}
\end{figure}

@@

In the output sum, module H has 2 types of beads representing "0" and  3 types of beads representing "1".
Then, module F is folded to F0 and F1 in Fig.~\ref{fig:formatters} according to the beads corresponding to 0 (H00, H11) and 1 (H01, H10, He1), respectively, to format beads representing a value.
\begin{figure}[h]
\begin{tabular}{c}
 \begin{minipage}{0.3\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H00_1.pdf}
 \end{minipage}

 \end{minipage}
 \end{tabular}
 \caption{All the 6 bricks of Module H. In zags, H always folds into the bottom-right brick Hn. }
 \label{fig:halfadders}
\end{figure}

\begin{figure}[h]
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.6\linewidth]{fig/svg/Fnt_1.pdf}
 \end{minipage}

 \end{minipage}
 \end{tabular}
 \caption{All the 4 bricks of Module F: The 2 bricks at the top are for zigs while the 2 bottom are for zags. }
 \label{fig:formatters}
\end{figure}

@@

Subsequently, module R and F are folded, and then module L folding starts (Fig.~\ref{fig:overflowex4}).
Since module L turns by binding to Turn Signal, it is necessary to give signal for turn in order to move to zag pass from here.
Therefore, let the top of folded module F be Turn Signal.
Normally, there are folded beads above module F, so module L can not bind to F's Turn Signal.
However, only when it overflows, above F becomes blank, so L turns by being folded there (Fig.~\ref{fig:overflowex5}).
The turned module L forms a Turn Signal at the end, and folding of zag pass starts.
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/svg/CounterEx13_1.pdf}
\caption{
Reach the left end with carry.
Even if transcript of module L sticks to Turn Signal, it can not bind because the distance is long.
}
\label{fig:overflowex1}
\end{figure}

@@

\begin{tabular}{c}
\ \\
\ \\
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Lbe_3.pdf}
 \end{minipage}

  \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Ltre_3.pdf}

 \end{minipage}
 \end{tabular}
 \caption{The paths of module L}
 \label{fig:leftturns}
\end{figure}

 \end{tabular}
 \caption{The paths of module R}
 \label{fig:rightturns}
\end{figure}

>>>>

%\titlerunning{}
\author{
Kohei Maruyama\inst{1}\thanks{Primary corresponding author (\email{k.maruyama@uec.ac.jp})} \and
Shinnosuke Seki\inst{1,2}\thanks{Secondary corresponding author (\email{s.seki@uec.ac.jp})}
}
\institute{
The University of Electro-Communications, 
%Graduate School of Informatics and Engineering, 
1-5-1 Chofugaoka, Chofu, Tokyo, 1828585, Japan \and
\'{E}cole Normale Superi\'{e}ure de Lyon, 46 all\'{e}e d'Italie, 69007, Lyon, France
}

@@

A fixed bit-width (finite) binary counter has been implemented as a proof-of-concept demonstration of oritatami model of cotranscriptional folding \cite{GeMeScSe2019}. 
As shown in Fig.~\ref{fig:rna_origami}, an RNA transcript folds upon itself while being transcribed (synthesized) from its corresponding DNA template strand. 
Geary et al. programmed a specific RNA rectangular tile structure into a DNA template in such a way that the corresponding RNA transcript \textit{folds cotranscriptionally} into the programmed tile structure highly likely \textit{in vitro} at room temperatures (\textit{RNA origami}) \cite{GearyRothemundAndersen2014}. 
An oritatami system folds a transcript of abstract molecules called \textit{beads} of finitely many types cotranscriptionally over the 2-dimensional triangular lattice cotranscriptionally according to a rule set that specifies which types of molecules are allowed to bind. 
The transcript of the binary counter in \cite{GeMeScSe2019} is of period 60 as $\textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1}{-}\textcircled{\scriptsize 2}{-} \cdots {-}\textcircled{\scriptsize 58}{-}\textcircled{\scriptsize 59}{-}\textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1} \cdots$ and its period is semantically divided into two half-adder (HA) modules $A = \textcircled{\scriptsize 0}{-}\textcircled{\scriptsize 1}{-} \cdots {-}\textcircled{\scriptsize 11}$ and $C = \textcircled{\scriptsize 30}{-}\textcircled{\scriptsize 31}{-} \cdots {-}\textcircled{\scriptsize 41}$ and two structural modules $B$ and $D$, which are sandwiched by half-adder modules.
While being folded cotranscriptionally in zigzags, HA modules increment the current count $i$ by 1, which is initialized on a linear \textit{seed} structure, alike the Evans' counter, whereas structural modules $B$ and $D$ align HA modules properly and also make a turn at an end of the count $i$; $B$ guides the transcript from a zig to a zag ($\hookrightarrow$) while $D$ does from a zag to a zig ($\hookleftarrow$). 
This counter was embedded as a component of an oritatami system to self-assemble an arbitrary finite portion of Heighway dragon fractal \cite{MasudaSekiUbukata2018}. 
Its applications are limited, however, by lack of mechanism to widen bit-width; its behavior is undefined when its count overflows. 
In this paper, we endow this counter, or more precisely, its structural module $B$, with capability to widen the count by 1 bit at every encounter with overflow.

@@

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.6\linewidth]{fig/svg/Fnt_1.pdf}
 \end{minipage}

 \end{minipage}
 \end{tabular}
 \caption{All the 4 bricks of module F: The 2 bricks at the top are for zigs while the 2 bottom are for zags. }
 \label{fig:formatters}
\end{figure}

@@

\subsection{General idea}
Between two consecutive overflows, the proposed system behaves in the same way as the finite binary counter of Geary et al. \cite{GeMeScSe2019}.
Its transcript folds in a zigzag manner macroscopically (downward in figures throughout this paper).
A zig, folding from right to left, increments the current value of the counter by 1.
The succeeding zag, folding from left to right, formats the incremented value for the sake of next zig and copies it downward.
Unlike the counter of Geary et al., when a zig encounters an overflow, it does not abort but extends the current value by 1 bit.

@@

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.3\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H00_1.pdf}
 \end{minipage}

 \end{minipage}
 \end{tabular}
 \caption{All the six bricks of module H: \texttt{H00}, \texttt{H01}, \texttt{He1}, \texttt{H10}, \texttt{H11}, and \texttt{Hn} from top left to bottom right.
In zags, H always folds into \texttt{Hn} while in zigs, it folds into one of the other five bricks.}
 \label{fig:halfadders}
\end{figure}

\begin{tabular}{c}
\ \\
\ \\
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/Ltre_3.pdf}
 \end{minipage}

  \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Lbe_3.pdf}
   \vspace*{18mm}

 \end{minipage}
 \end{tabular}
 \caption{All the five bricks of module L: \texttt{Lt}, \texttt{Lbn}, \texttt{Lcrn}, \texttt{Lcre}, and \texttt{Lbe} from top left to bottom right.
In zigs, L folds into either \texttt{Lt} or \texttt{Lbn} depending on where it starts, until the transcript reaches the left end, where L folds either into \texttt{Lcrn} if the current value has not been overflowed, or into \texttt{Lbe} at an overflow.
In the case of overflow, the next L folds into \texttt{Lcre}.
In zags, L always folds into \texttt{Lbn}.}
 \label{fig:leftturns}
\end{figure}

 \end{tabular}
 \caption{All the three bricks of module R: \texttt{Rt}, \texttt{Rb}, and \texttt{Rcr}.
 In zigs, R folds into \texttt{Rt} or \texttt{Rb}.
 In zags, R always folds into \texttt{Rb} until the transcript reaches the right end, where R folds into \texttt{Rcr}.}
 \label{fig:rightturns}
\end{figure}

\paragraph{Seed and Encoding.}
The initial counter value is encoded as $b_{k-1}b_{k-2} \cdots b_1b_0$ in binary on the seed in the following format
\begin{equation} \label{eq:zagencoding}
64{-}65{-}66{-}\left( \prod^0_{i = k-1} \bigl(  w_{Hn} w_{Rb} w_{Fb_i} w_{Lbn} \bigr) \right) w_{Hn}
\end{equation}
where $w_{Hn}, w_{Rb}, w_{Fb_i}, w_{Lbn}$ are sequences of bead types exposed downward by modules H, R, F, L when they fold into bricks Hn, Rb, F$b_i$, Lbn, respectively, which can be found in Figs.~\ref{fig:halfadders}, \ref{fig:rightturns}, \ref{fig:formatters}, and \ref{fig:leftturns}.
The seed for $k = 1$ and $b_0 = 0$ is colored in purple in Fig.~\ref{fig:counter1stzig}.
%$w_{Hn} = 67{-}76{-}77{-}78{-}79{-}88{-}89{-}90{-}91{-}96$ \\
%$w_{Rb} = 97{-}102{-}103{-}108{-}109{-}114{-}115{-}120{-}121{-}126{-}127{-}132$\\
%$w_{Lbn} = 31{-}36{-}37{-}42{-}43{-}48{-}49{-}54{-}55{-}64{-}65{-}66$

\subsubsection{In the zig ($\leftarrow$)} 
Here, the counter lets instances of module H increment its current value, which is encoded in the format \eqref{eq:zagencoding} below the seed or previous zag.
When the current value is $k$ bits in width, a subsequence $(FLHR)^kF$ of the transcript folds into the zig.
All the bricks which modules take in a zig are of height 3 (see Fig.~\ref{fig:formatters},  \ref{fig:halfadders}, \ref{fig:leftturns}, and \ref{fig:rightturns}) so that the zig turns out to be a linear structure of height 3.
Carry is propagated throughout the zig as a height for modules to start.
All modules but H end at the same height as it began, that is, do not change carry.
Module H can take the five bricks \texttt{H00}, \texttt{H01}, \texttt{H10}, \texttt{H11}, and \texttt{He1} shown in Fig.~\ref{fig:halfadders} in zigs but nothing else.
This is because the whole system is designed in such a way that module H encounters only the five environments.
Until the count overflows, module H encounters only four of them, which encode input 0 as $w_{F0}$ or 1 as $w_{F1}$ and carry or no-carry as the height for the module to start folding (bottom or top), where it takes \texttt{H00}, \texttt{H01}, \texttt{H10}, and \texttt{H11}, respectively, as shown in Fig.~\ref{fig:halfadders} (\texttt{H}$xc$ is folded when the input is $x$ and the carry is given if $c=1$ or not otherwise).

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/svg/CounterEx5_1.pdf}
\caption{
The first zig pass.
The seed is encoded with "0" same as bottom of F0(
\textcircled{\scriptsize 1}{-}\textcircled{\scriptsize 10}{-}\textcircled{\scriptsize 11}{-}\textcircled{\scriptsize 12}{-}\textcircled{\scriptsize 13}{-}\textcircled{\scriptsize 22}{-}\textcircled{\scriptsize 23}{-}\textcircled{\scriptsize 24}{-}\textcircled{\scriptsize 25}{-}\textcircled{\scriptsize 30}),
 and the counter increments it by giving a carry.
Module H outputs "1" as a sum and cancels the carry.
}
\label{fig:counter1stzig}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/svg/CounterEx13_1.pdf}
\caption{
Reach the left end with carry.
Even if transcript of module L sticks to Turn Signal, it can not bind because the distance is long.
}
\label{fig:overflowex1}
\end{figure}

Let us see how the subsequence $(FLHR)^kF$ folds into a zig; for $k=1$ and the current value 0 see Fig.~\ref{fig:counter1stzig}.
The zig is initially fed with carry by being forced to start folding at the bottom and the carry is propagated by modules F, L, and R in the way just explained.
The first instance of H is thus fed with carry and folds into \texttt{H01} if the bit encoded above is 0, as illustrated in Fig.~\ref{fig:counter1stzig}, or \texttt{H11} if the bit is rather 1.
\texttt{H01} ends at the top, corresponding to no-carry.
This absence of carry propagates through the succeeding modules leftward.
As a result, the zig ends, or more precisely an instance of F ends folding at the bottom if the current value is overflowed (Fig.~\ref{fig:overflowex1}), or at the top otherwise. (Fig.~\ref{fig:counter1stzig}).
An instance of L is to be transcribed next.
It folds either into \texttt{Lcrn} for (normal) carriage-return unless the current value is overflowed, or into \texttt{Lbe} at an overflow.
% Lbeの後の話 Lcreまで

Module H read from the row above the value encoded into folding in the module F during the previous zag-phase, and fold into a shape H00, H10, H01, H11.
H$xc$ is corresponding to the case where $x$ is the bit read in the row above and $c$ is the carry. 
In the zig pass module R, F, and L just propagate the carry value to the next.

@@

\subsubsection{In the zag ($\rightarrow$)}
This part, instances of module F copy the current value, which incremented in the previous zig and encoded at below the zig.
Module F is always folded under module H in zags and F takes the two bricks F0 and F1 shown in Fig.~\ref{fig:formatters} which depending on a brick of module H.
Other modules H,R, and L take bricks which do nothing Hn, Rb, and Lbc, respectively, shown in Fig.~\ref{fig:halfadders}, \ref{fig:rightturns}, \ref{fig:leftturns}.

a

Module F read form the row above H$xc$, and fold into a shape F0, F1.
F$y$ is corresponding to the case where $y = (x + c)$ mod $2$.
There are no carry propagation and the module L, H, R fold into Lbc, Hn, Rb in this pass.

@@

@@

In the output sum, module H has 2 types of beads representing "0" and  3 types of beads representing "1".
Then, module F is folded to F0 and F1 in Fig.~\ref{fig:formatters} according to the beads corresponding to 0 (H00, H11) and 1 (H01, H10, He1), respectively, to format beads representing a value.

@@

Subsequently, module R and F are folded, and then module L folding starts (Fig.~\ref{fig:overflowex4}).
Since module L turns by binding to Turn Signal, it is necessary to give signal for turn in order to move to zag pass from here.
Therefore, let the top of folded module F be Turn Signal.
Normally, there are folded beads above module F, so module L can not bind to F's Turn Signal.
However, only when it overflows, above F becomes blank, so L turns by being folded there (Fig.~\ref{fig:overflowex5}).
The turned module L forms a Turn Signal at the end, and folding of zag pass starts.

@@
